An efficient acyclic contact planner for multiped robots

### 2015-12-31 body/abstract.tex
cosmetics, added title, extended conclusion

We address the issue of automatically planning discrete sequences of contacts
that characterize a multiped motion in an arbitrary cluttered environment.
To overcome the inherent combinatorial difficulty of this problem, we separate it
in two sub-problems: first, planning a guide path for the root of the robot and
then, generating relevant contacts along this path.
We propose theoretical contributions for each sub-problem, and implement them
as a complete contact planner. 
First, we propose a geometrical characterization of a guide trajectory, the ``reachability condition", which guarantees that a
guide can be mapped into the contact manifold of the robot. With this condition, asserting the``true feasibility" 
of a guide path boils down to performing collision tests, thus avoiding the expensive explicit computation of balanced contact configurations, as opposed to previous approaches.
Then, we propose a sequential algorithm to extend a truly feasible path into a sequence of contacts. For each limb, an off-line sampling approach,
independent from the environment, is run once and for all, and used at run time to provide efficiently candidate configurations for contact.
The best candidate is chosen according to user-defined heuristics. We propose two such heuristics, used in all our experiments, and 
designed for generic robotic applications: a manipulatibily-based heuristic, EFORT, and a robustness criterion regarding the balance of the robot.
These contributions are implemented within the open-source software Hpp. Our framework allows us to generate complex multi-contact sequences for a large
variety of robots, including HRP-2, Hyq, and a dexterous hand, with interactive performances.

>>>>

We address the issue of automatically planning discrete sequences of contacts
that characterize a multiped motion in an arbitrary cluttered environment.
To overcome the inherent combinatorial difficulty of this problem, we separate it
in two sub-problems: first, planning a guide path for the root of the robot and
then, generating relevant contacts along this path.
We propose theoretical contributions for each sub-problem, and implement them
as a complete contact planner. 
First, we propose a geometrical characterization of a guide trajectory, the ``reachability condition", which guarantees that a
guide can be mapped into the contact manifold of the robot. With this condition, asserting the``true feasibility" 
of a guide path boils down to performing collision tests, thus avoiding the expensive explicit computation of balanced contact configurations, as opposed to previous approaches.
Then, we propose a sequential algorithm to extend a truly feasible path into a sequence of contacts. For each limb, an off-line sampling approach,
independent from the environment, is run once and for all, and used at run time to provide efficiently candidate configurations for contact.
The best candidate is chosen according to user-defined heuristics. We propose two such heuristics, used in all our experiments, and 
designed for generic robotic applications: a manipulability-based heuristic, EFORT, and a robustness criterion regarding the balance of the robot.
These contributions are implemented within the open-source software Hpp. Our framework allows us to generate complex multi-contact sequences for a large
variety of robots, including HRP-2, HyQ, and a dexterous hand, with interactive performances.

### 2015-12-31 body/appendix_algo.tex
typos
cosmetics, added title, extended conclusion

\section{Algorithms for discretizing of a path}
\label{app:contact}

@@

At each step, $g$ is called with the previous state as a parameter, as well
as a new root placement. $g$ returns a new contact configuration, if it suceeded
in computing a configuration with only one contact switch occuring.
If $g$ failed in achieving this, the method \textsc{RepositionContacts} is called.
It repositions one end effector (either a free limb, or the oldest active contact) towards a new contact position if possible.
This repositionning allows to increase the odds that the contact can be maintained at the next step.

>>>>

\section{Algorithms for the discretization of a path}
\label{app:contact}

@@

At each step, $g$ is called with the previous state as a parameter, as well
as a new root placement. $g$ returns a new contact configuration, if it succeeded
in computing a configuration with only one contact switch occurring.
If $g$ failed in achieving this, the method \textsc{RepositionContacts} is called.
It repositions one end effector (either a free limb, or the oldest active contact) towards a new contact position if possible.
This repositioning allows to increase the odds that the contact can be maintained at the next step.

### 2015-12-31 body/conclusion.tex
cosmetics, added title, extended conclusion

One direction for future work is to focus on a more accurate formulation of $C_{reach}$ to address 
this limitation.
Our main objective is of course to generate the complete interpolation between the contact sequences.
We have already been able to achieve this for HRP-2 for some of the computed sequences~\citep{Carpentier2016}, but we admit that additional
work is required on the planning side to obtain a seamless workflow. To achieve this, we are currently working on the notion of transition certificate, i.e. formulating
conditions that guarantee that the interpolation between two contact configurations is dynamically feasible. We are also
working on new heuristics more closely related to the dynamics of the systems (for instance, choose the contacts that minimize the jerk between two contact configurations).
\adnote{The 'future work' part seems a bit poor to me}

>>>>

One direction for future work is to focus on a more accurate formulation of $C_{reach}$ to address 
this limitation.
Our main objective is of course to generate the complete interpolation between the contact sequences.
We have already been able to achieve this for HRP-2 for some of the computed sequences~\citep{Carpentier2016}, but we admit that additional
work is required on the planning side to obtain a seamless workflow. To achieve this, we are currently working on the notion of transition certificate, i.e. formulating
conditions that guarantee that the interpolation between two contact configurations is dynamically feasible.
To achieve this, our objective is to propose a geometric quantification of both the kinematic and dynamic constraints of the problem, expressed
at the center of mass of the robot.
 We are also
working on new heuristics more closely related to the dynamics of the systems (for instance, choose the contacts that minimize the jerk between two contact configurations).
These two points illustrate the need to reduce the distance between the planning and control aspects of multiped locomotion. We believe that one way 
to bridge the gap is to introduce robustness while addressing these two issues.

### 2015-12-31 body/intro.tex
cosmetics, added title, extended conclusion

Most multipedal locomotion systems focus on cyclic walking gaits \citep{Kajita03a}. However, executing
this behavior on cluttered environments is dangerous, if not impossible.
In an analysis of their participation to the Darpa Robotics Challenge, \citeauthor{atkensondarpa}
noted: ``Except for egress, no robots in the DRC Finals used
the  stair  railings  or  any  form  of  bracing.   Even  drunk  people  are  smart  enough  to  use  nearby  supports.
Full body locomotion (handholds,  bracing,  leaning against a wall or obstacles) should be easier than our
current high performance minimum contact locomotion approaches."

@@

As suggested by \citeauthor{Bouyarmane2009}, we believe that these two tasks can be treated sequentially within a global planner, while preserving the completeness of the search.
We go further and claim that this can be done at a much smaller cost, provided that necessary and sufficient conditions for contact feasibility can be formulated for the robot trajectory.
This paper presents a geometrical representation of these conditions, and a concrete implementation of such a decoupled planner.
This implementation results from a trade-off between a necessary and a sufficient condition for feasibility, allowing us to find solutions extremely fast.
while preserving a success rate greater than TODO \% in the demonstrated scenarios.

>>>>

Most multipedal locomotion systems focus on cyclic walking gaits \citep{Kajita03a}. However, executing
this behavior on cluttered environments is dangerous, if not impossible.
In an analysis of their participation to the Darpa Robotics Challenge, \citeauthor{atkensondarpa}
noted: ``Except for egress, no robots in the DRC Finals used
the  stair  railings  or  any  form  of  bracing.   Even  drunk  people  are  smart  enough  to  use  nearby  supports.
Full body locomotion (hand holds,  bracing,  leaning against a wall or obstacles) should be easier than our
current high performance minimum contact locomotion approaches."

@@

As suggested by \citeauthor{Bouyarmane2009}, we believe that these two tasks can be treated sequentially within a global planner, while preserving the completeness of the search.
We go further and claim that this can be done at a much smaller cost, provided that necessary and sufficient conditions for contact feasibility can be formulated for the robot trajectory.
This paper presents a geometrical representation of these conditions, and a concrete implementation of such a decoupled planner.
This implementation results from a trade-off between a necessary and a sufficient condition for feasibility, allowing us to find solutions extremely fast.
while preserving a high success rate in the demonstrated scenarios.

### 2015-12-31 body/overview.tex
cosmetics, added title, extended conclusion

From a given start configuration, the planner proceeds in an iterative fashion along the discretized path: given a new root placement, an inverse-kinematics solver 
is used to maintain the contacts that were previously existing (TODO Figure). Often, these contacts cannot be maintained, because of joint limits or collisions (TODO Figure).
In such a case the contacts are broken. Conversely, new contacts are created to ensure the static equilibrium of the robot (TODO FIGURE).
The algorithm is designed so that only one contact can be created or broken between two successive configurations. This ensures that the interpolation between two configurations is achievable.
Details are presented in Section~\ref{sec:contact}.

>>>>

From a given start configuration, the planner proceeds in an iterative fashion along the discretized path: given a new root placement, an inverse-kinematics solver 
is used to maintain the contacts that were previously existing. Often, these contacts cannot be maintained, because of joint limits or collisions.
In such a case the contacts are broken. Conversely, new contacts are created to ensure the static equilibrium of the robot.
The algorithm is designed so that only one contact can be created or broken between two successive configurations. This ensures that the interpolation between two configurations is achievable.
Details are presented in Section~\ref{sec:contact}.

### 2015-12-31 body/rbprm.tex
cosmetics, added title, extended conclusion

Therefore we can approximate $C_{reach}$ with a set $C_{reach} \subset C_{reach}^1$ with the reachability condition defined as: 
\begin{equation}
C_{reach}^1 = \{ \mathbf{q}^0 : W(\mathbf{q}^{0}) \cap O \neq \emptyset \wedge W^0(\mathbf{q}^{0}) \cap O = \emptyset \} % \\ }
 %~ & \wedge & A_{trunk}(\mathbf{q}^{root}) \cap W = \emptyset \}
\end{equation}
It is straighforward to prove that  $C_{reach} \subset C_{reach}^1$ (by construction of the included set). 
This inclusion is very important: it directly implies that any motion-planning algorithm with a guaranty of completeness in $SE(3) \times \mathbb{R}^n$ is complete  in $C_{reach} \times \mathbb{R}^n$. This is a strong reduction of the search space, which can be directly applied to any existing method.

@@

However, between $W^0$ and $B^{max}$, a trade-off can be found between a necessary and a sufficient condition. We define $W^0_s$ as the volume $W^0$ subject to a scaling transformation by a factor $s \in \mathbb{R}^+$. 
%
We then consider the spaces $C_{reach}^s$
 \begin{equation}
C_{reach}^s = \{ \mathbf{q}^0 : W(\mathbf{q}^{0}) \cap O \neq \emptyset \wedge W^0_s(\mathbf{q}^{0}) \cap O = \emptyset \} % \\ }
 %~ & \wedge & A_{trunk}(\mathbf{q}^{root}) \cap W = \emptyset \}
\end{equation}
%
The higher $s$, the closer this ``reachability condition'' is to being sufficient, and if $s=1$, the planner is complete. The parametrization of $s$ allows to find a trade-off between these two desirable properties.
As an example, Figure~\ref{fig:hyq_roms} presents the volumes $W$ computed for the HyQ robot.
In Appendix~\ref{app:rom}, we give a generic method to compute these volumes, with the example of HRP-2.
%~ Section~\ref{sec:results} shows that in practice, it is easy to adjust $s$ to keep most of the interesting guides without introducing incorrect guides.

\begin{figure}
  \centering
  \includegraphics[width=0.95\linewidth]{figures/hyq_roms}
  \caption{
           $W$ volumes for HyQ. The green shapes represent the reachable workspace $W^k$ of each limb. The red shape is $W^0$, a scaling of the convex hull
           of the torso of the robot.}
		   \label{fig:hyq_roms}
\end{figure}

\subsection{Computing the guide trajectory in $C_{reach}^s$}
Once a value of $s$ has been fixed, any sampling-based motion planner can be used to plan a trajectory in $C_{reach}^s$. 
Indeed, contrary to $C_{Contact}$,  $C_{reach}^s$ has a non zero measure in the configuration space $C$. Therefore a standard uniform sampling approach
can work, in spite of a high rejection rate. 
Thus, the only significant change regarding a classical planner is to replace the classical collision checking method with a test of appartenance to $C_{reach}^s$ when verifying
that drawn configurations and associated local paths are valid.

@@

Thanks to this small set of modifications, the high dimensional truly feasible path planning problem is reduced to a standard geometry and collision checking problem, in a much lower dimension (6 for Hyq, 8 for HRP-2 that has joints in the torso). In so doing we make the problem tractable for sampling-based approaches, with extremely efficient computation times.

>>>>

Therefore we can approximate $C_{reach}$ with a set $C_{reach} \subset C_{reach}^1$ with the reachability condition defined as: 
\begin{equation}
C_{reach}^1 = \{ \mathbf{q}^0 : W(\mathbf{q}^{0}) \cap O \neq \emptyset \wedge W^0(\mathbf{q}^{0}) \cap O = \emptyset \} % \\ }
 %~ & \wedge & A_{trunk}(\mathbf{q}^{root}) \cap W = \emptyset \}
\end{equation}
It is straightforward to prove that  $C_{reach} \subset C_{reach}^1$ (by construction of the included set). 
This inclusion is very important: it directly implies that any motion-planning algorithm with a guaranty of completeness in $SE(3) \times \mathbb{R}^n$ is complete  in $C_{reach} \times \mathbb{R}^n$. This is a strong reduction of the search space, which can be directly applied to any existing method.

@@

However, between $W^0$ and $B^{max}$, a trade-off can be found between a necessary and a sufficient condition. We define $W^0_s$ as the volume $W^0$ subject to a scaling transformation by a factor $s \in \mathbb{R}^+$. 
%
We then consider the spaces $C_{reach}^s$
 \begin{equation}
C_{reach}^s = \{ \mathbf{q}^0 : W(\mathbf{q}^{0}) \cap O \neq \emptyset \wedge W^0_s(\mathbf{q}^{0}) \cap O = \emptyset \} % \\ }
 %~ & \wedge & A_{trunk}(\mathbf{q}^{root}) \cap W = \emptyset \}
\end{equation}
%
The higher $s$, the closer this ``reachability condition'' is to being sufficient, and if $s=1$, the planner is complete. The parametrization of $s$ allows to find a trade-off between these two desirable properties.
As an example, Figure~\ref{fig:HyQ_roms} presents the volumes $W$ computed for the HyQ robot.
In Appendix~\ref{app:rom}, we give a generic method to compute these volumes, with the example of HRP-2.
%~ Section~\ref{sec:results} shows that in practice, it is easy to adjust $s$ to keep most of the interesting guides without introducing incorrect guides.

\begin{figure}
  \centering
  \includegraphics[width=0.95\linewidth]{figures/HyQ_roms}
  \caption{
           $W$ volumes for HyQ. The green shapes represent the reachable workspace $W^k$ of each limb. The red shape is $W^0$, a scaling of the convex hull
           of the torso of the robot.}
		   \label{fig:HyQ_roms}
\end{figure}

\subsection{Computing the guide trajectory in $C_{reach}^s$}
Once a value of $s$ has been fixed, any sampling-based motion planner can be used to plan a trajectory in $C_{reach}^s$. 
Indeed, contrary to $C_{Contact}$,  $C_{reach}^s$ has a non zero measure in the configuration space $C$. Therefore a standard uniform sampling approach
can work, in spite of a high rejection rate. 
Thus, the only significant change regarding a classical planner is to replace the classical collision checking method with a test of belonging to $C_{reach}^s$ when verifying
that drawn configurations and associated local paths are valid.

@@

Thanks to this small set of modifications, the high dimensional truly feasible path planning problem is reduced to a standard geometry and collision checking problem, in a much lower dimension (6 for HyQ, 8 for HRP-2 that has joints in the torso). In so doing we make the problem tractable for sampling-based approaches, with extremely efficient computation times.

### 2015-12-31 body/results.tex
cosmetics, added title, extended conclusion

%~ \subsubsection{HRP-2 -- Truck egress (Figure~\ref{fig:TODO}):}
%~ HRP-2 stands half seated in a truck, and \replaced{is required to}{it must} \replaced{get out}{exit} through the front window.

%~ \begin{figure*}
  %~ \centering
  %~ \includegraphics[width=0.5\linewidth]{figures/standing}
  %~ \caption{
           %~ HRP-2 in the standing scenario. }
		   %~ \label{fig:standing}
%~ \end{figure*}

%~ 
%~ \noindent\textbf{Contacts involved:} All (both feet and hands).
%~ 
%~ \noindent\textbf{Heuristics:} $h_w$ \deleted{is used} for the feet, $h_{EFORT}$ \deleted{is used} for the hands.
%~ 
%~ \noindent\textbf{Observations:} This scenario demonstrates the ability of the planner to compute a path in an extremely-constrained environment. In this case,
%~ the scenario is not always real time (TODO or is it ?): while the planner is often able to compute a guide path rapidly, in the worst case up to one minute can be spent
%~ in the planning.

@@

\subsubsection{HyQ -- Obstacle race (Figure~\ref{fig:hyq_bridge} and~\ref{fig:hyq_obs}):}
In this long scene, HyQ is first required to cross a 55-cm large hole; then, to cross a narrow "bridge",  only 25-cm large.

\begin{figure}
  \centering
  \includegraphics[width=0.4\linewidth]{figures/hyq_bridge}
  \caption{
           HyQ crossing a narrow bridge. }
		   \label{fig:hyq_bridge}
\end{figure}

\begin{figure*}
  \centering
  \includegraphics[width=0.5\linewidth]{figures/hyq_obs}
  \caption{
           Crossing a hole contact sequence for HyQ ($b_0 > 4$). }
		   \label{fig:hyq_obs}
\end{figure*}

@@

\noindent\textbf{Observations:} Despite the apparent simplicity of the scene, this scenario is the hardest for our planner.
While finding a guide path above the hole is easy for the planner, finding a sequence of contacts that allows for equilibrium is not trivial.
Second, the narrow bridge is hard both for the planner and the contact generator: to make sure that equilibrium is preserved along the traversal,
the bridge must be approached with the right angle.
This can is illustrated in Figure~\ref{fig:hyq_obs}, where several feet rearrangements are required to cross both obstacles (although the video shows this best).
The planner however succeeds in finding a feasible sequence in the end, again with interactive computation times.

@@

\begin{table*}
\centering
\small
\begin{tabular}{ l | >{\centering\arraybackslash}m{57pt} | >{\centering\arraybackslash}m{70pt} | >{\centering\arraybackslash}m{75pt} | >{\centering\arraybackslash}m{77pt} | >{\centering\arraybackslash}m{85pt} | c}
  &  Complete guide generation & Static equilibrium & Collision & ik  & Total generation time & 1 step\\
 \hline
   Stairs & 6 -- \textbf{13} --  24 & 13 --  \textbf{32} -- 329   & 1 --  \textbf{4} -- 38 & 26 --  \textbf{127} -- 1345 & 92 --  \textbf{261} -- 2174 & \textbf{14} \\
   Standing & 4 -- \textbf{451} --  985 & 27 --  \textbf{144} -- 338   & 2 --  \textbf{12} -- 37 & 144 --  \textbf{1046} -- 2374 & 310 --  \textbf{1622} -- 3429 & \textbf{66} \\
   Rubble & 3 -- \textbf{524} --  993 & 242 --  \textbf{511} -- 3480   & 233 --  \textbf{505} -- 4564 & 180 --  \textbf{414} -- 3518 & 1400 --  \textbf{3058} -- 13126 & \textbf{43} \\
   Race & 1 -- \textbf{501} --  995 & 266 --  \textbf{449} -- 3956   & 824 --  \textbf{1061} -- 2130 & 666 --  \textbf{874} -- 1613 & 2530 --  \textbf{3722} -- 10030 & \textbf{33} \\
   %Standing up & $70$ & 23 & 48\\
 \end{tabular}
\caption{minimum, \textbf{average} and worst time (in ms) spent in the generation process for each scenario and each critical part of the generation process. The last
column indicates the average time necessary to compute one contact transition.}
\label{tab:requestime}
\quad
\end{table*}

In most scenarios, one can observe that the average computation time for one single step is largely below the second,
thus allowing to consider interactive applications. 
Considering the repartition of the computation time, it appears that most of the time is spent performing inverse kinematics.
This is not surprising considering the number of calls to the methods: IK is used intensively to maintain contact continuity between two postures; 
it is also applied every time a new candidate needs to be evaluated. However regarding state-of-the-art IK solutions, we believe there is room
for improving the performances of the algorithm.

\subsubsection{Success rates}
Table~\ref{tab:requestpercent} summarizes the success rates obtained for each scenario.
\begin{table*}
\centering
\begin{tabular}{ l | >{\centering\arraybackslash}m{65pt} | >{\centering\arraybackslash}m{65pt} | >{\centering\arraybackslash}m{65pt} | >{\centering\arraybackslash}m{65pt} | c}
  &  Path planning & True feasibility & Kinematic failure & Equilibrium failure \\
 \hline
   Steep stairs & 100\%  & 99.5\% & 0.11\% & 0.39\% \\
   Standing up & 68\% & 88\% & 6\% & 6\% \\
   %~ Truck egress & \% & \% & \% &  \% \\
   Rubble & 74\% & 97.87\% & 0.13\% & 2\% \\
   Obstacle race & 58\% & 95.7\% & 1.8\% & 2.5\% \\
 \end{tabular}
\caption{Success rates for each scenario. The first value indicates the percentage of successfull complete contact plannings for 1000 tests; The second value
indicates the percentage of truly feasible root configurations: considering each limb individually, indicates the percentage of root placements of the guide trajectory that
led to a feasible contact. Kinematic failure is the percentage of contact generations that failed because no collision-free candidate was found. Equilibrium failure is the percentage of contact
generations that failed because no candidate respected the static equilibrium condition.}
\label{tab:requestpercent}
\quad
\end{table*}

We observe that, as expected, our planner does not succeed systematically, because of the approximations made in the implementation
of our theoretical formulation. 
From a pragmatic point of view, regarding the computation times, we claim that our approach is the best current compromise between completeness and efficiency:
Indeed, the advantage of the framework is that when the contact generation fails, it does so rapidly, which allows us to rapidly re-plan a new contact sequence with 
a reasonable chance of success.

>>>>

@@

\subsubsection{HyQ -- Obstacle race (Figure~\ref{fig:HyQ_bridge} and~\ref{fig:HyQ_obs}):}
In this long scene, HyQ is first required to cross a 55-cm large hole; then, to cross a narrow "bridge",  only 25-cm large.

\begin{figure}
  \centering
  \includegraphics[width=0.4\linewidth]{figures/HyQ_bridge}
  \caption{
           HyQ crossing a narrow bridge. }
		   \label{fig:HyQ_bridge}
\end{figure}

\begin{figure*}
  \centering
  \includegraphics[width=0.5\linewidth]{figures/HyQ_obs}
  \caption{
           Crossing a hole contact sequence for HyQ ($b_0 > 4$). }
		   \label{fig:HyQ_obs}
\end{figure*}

@@

\noindent\textbf{Observations:} Despite the apparent simplicity of the scene, this scenario is the hardest for our planner.
While finding a guide path above the hole is easy for the planner, finding a sequence of contacts that allows for equilibrium is not trivial.
Second, the narrow bridge is hard both for the planner and the contact generator: to make sure that equilibrium is preserved along the traversal,
the bridge must be approached with the right angle.
This can is illustrated in Figure~\ref{fig:HyQ_obs}, where several feet rearrangements are required to cross both obstacles (although the video shows this best).
The planner however succeeds in finding a feasible sequence in the end, again with interactive computation times.

@@

\begin{table*}
\centering
\footnotesize
\begin{tabular}{ >{\centering\arraybackslash}m{37pt} | >{\centering\arraybackslash}m{57pt} | >{\centering\arraybackslash}m{65pt} | >{\centering\arraybackslash}m{70pt} | >{\centering\arraybackslash}m{73pt} | >{\centering\arraybackslash}m{80pt} | >{\centering\arraybackslash}m{10pt}}
  Scenario (nb steps) &  Complete guide generation & Static equilibrium & Collision & Inverse Kinematics  & Total generation time & Time per step\\
 \hline
   Stairs (18) & 6 -- \textbf{13} --  24 & 13 --  \textbf{32} -- 329   & 1 --  \textbf{4} -- 38 & 26 --  \textbf{127} -- 1345 & 92 --  \textbf{261} -- 2174 & \textbf{14} \\
   Standing (24)& 4 -- \textbf{451} --  985 & 27 --  \textbf{144} -- 338   & 2 --  \textbf{12} -- 37 & 144 --  \textbf{1046} -- 2374 & 310 --  \textbf{1622} -- 3429 & \textbf{66}  \\
   Rubble (71)& 3 -- \textbf{524} --  993 & 242 --  \textbf{511} -- 3480   & 233 --  \textbf{505} -- 4564 & 180 --  \textbf{414} -- 3518 & 1400 --  \textbf{3058} -- 13126 & \textbf{43} \\
   Race (112)& 1 -- \textbf{501} --  995 & 266 --  \textbf{449} -- 3956   & 824 --  \textbf{1061} -- 2130 & 666 --  \textbf{874} -- 1613 & 2530 --  \textbf{3722} -- 10030 & \textbf{33} \\
   %Standing up & $70$ & 23 & 48\\
 \end{tabular}
\caption{minimum, \textbf{average} and worst time (in ms) spent in the generation process for each scenario and each critical part of the generation process. The last
column indicates the average time necessary to compute one contact transition.}
\label{tab:requestime}
\quad
\end{table*}

Considering the repartition of the computation time, for HRP-2, most of the time is spent performing inverse kinematics.
This is not surprising considering the number of calls to the methods: IK projection is used intensively to maintain contact continuity between two postures; 
it is also applied every time a new candidate needs to be evaluated. In particular for the standing up scenario,
the kinematic constraints are really strong. Therefore a large number of projections fail because of joint limits violation.

For HyQ, there is a more regular repartition of the performances for the rubble scenario. For the obstacle scenario,
we observe a large increase of the time spent performing collision checking and inverse kinematics. This is explained
by the complexity of crossing the bridge: the robot has to stand high and create contacts close to each other because
the bridge is narrow. In many cases this results in collisions, thus a large number of contacts have to be evaluated.

In all scenarios, one can observe that the average computation time for one single step is largely below the second,
thus allowing to consider interactive applications.

\subsubsection{Success rates}
Table~\ref{tab:requestpercent} summarizes the success rates obtained for each scenario.
\begin{table*}
\centering
\begin{tabular}{ l | >{\centering\arraybackslash}m{65pt} | >{\centering\arraybackslash}m{65pt} | >{\centering\arraybackslash}m{65pt} | >{\centering\arraybackslash}m{65pt} | c}
  &  Path planning & True feasibility & Kinematic failure & Equilibrium failure \\
 \hline
   Steep stairs & 100\%  & 99.5\% & 0.11\% & 0.39\% \\
   Standing up & 68\% & 88\% & 6\% & 6\% \\
   %~ Truck egress & \% & \% & \% &  \% \\
   Rubble & 74\% & 97.87\% & 0.13\% & 2\% \\
   Obstacle race & 58\% & 95.7\% & 1.8\% & 2.5\% \\
 \end{tabular}
\caption{Success rates for each scenario. The first value indicates the percentage of successful complete contact plannings for 1000 tests; The second value
indicates the percentage of truly feasible root configurations: considering each limb individually, indicates the percentage of root placements of the guide trajectory that
led to a feasible contact. Kinematic failure is the percentage of contact generations that failed because no collision-free candidate was found. Equilibrium failure is the percentage of contact
generations that failed because no candidate respected the static equilibrium condition.}
\label{tab:requestpercent}
\quad
\end{table*}

We observe that, as expected, our planner does not succeed systematically, because of the approximations made in the implementation
of our theoretical formulation. 
From a pragmatic point of view, regarding the computation times, we claim that our approach is the best current compromise between completeness and efficiency.
Indeed, the advantage of the framework is that when the contact generation fails, it does so rapidly, which allows us to rapidly re-plan a new contact sequence with a reasonable chance of success.
The most efficient (and immediate) approach is probably to launch in parallel several instances of the planner for a given problem (our current implementation is single threaded) and to apply any successful result.

### 2015-12-31 main.tex
revised beg intro
intro + overview no fig
debut planner
finished contact generation now heuristics
robust equilibrium
corrections andrea
appendix
Add my comments to abstract
Andrea's first complete review
cosmetics, added title, extended conclusion

\setcounter{secnumdepth}{3}
\usepackage{moreverb,url}

@@

\title{A demonstration of the \LaTeXe\ class file for
\itshape{SAGE Publications}}

\author{Steve Tonneau\affilnum{1}, Andrea Del Prete\affilnum{1}, Julien Pettr\'e\affilnum{2}, Franck Multon\affilnum{3}, Nicolas Mansard\affilnum{1}}

\affiliation{\affilnum{1}LAAS-CNRS, Toulouse, France\\
\affilnum{2}Inria, Rennes, France \\
\affilnum{3}Universit\'e Rennes 2, France}

@@

\keywords{Class file, \LaTeXe, \textit{SAGE Publications}}

@@

\section{Introduction}
\input{body/intro}

\end{document}

>>>>

\setcounter{secnumdepth}{3}
\usepackage{moreverb,url}
\usepackage{makeidx}         % allows index generation
\usepackage[percent]{overpic}
\usepackage{array} 
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{listings}
\usepackage{color}
\usepackage{bm}

\input{./math_commands.tex}

\newcommand{\commentt}[2]{\textcolor{#1}{\textbf{\textit{[#2]}}}} 	% make sure no comments left in the paper
%\newcommand{\commentt}[2]{}							% using this line rather than the previous one you remove all comments
\newcommand{\adnote}[1]{\commentt{blue}{AD: #1}}
\newcommand{\nmnote}[1]{\commentt{OliveGreen}{NM: #1}}

% Another method to track changes
% Examples of usage:
% - This is \added[id=per,remark={we need this}]{new} text.
% - This is \deleted[id=per,remark=obsolete]{unnecessary}text.
% - This is \replaced[id=per]{nice}{bad} text.
% To print the list of change use: \listofchanges
\usepackage{changes}	% use this for the working version
%\usepackage[final]{changes} % use this for the final version
\definechangesauthor[name={Andrea Del Prete}, color=orange]{adp}

@@

\title{An efficient acyclic contact planner for multiped robots}

\author{Steve Tonneau\affilnum{1}, Andrea Del Prete\affilnum{1}, Julien Pettr\'e\affilnum{2}, Nicolas Mansard\affilnum{1}}
 %~ Franck Multon\affilnum{3},
\affiliation{\affilnum{1}LAAS-CNRS, Toulouse, France\\
\affilnum{2}Inria, Rennes, France } % \\
%~ \affilnum{3}Universit\'e Rennes 2, France}

@@

\keywords{Motion planning, Contact planning, Humanoid Robot, Legged Locomotion, Rough Terrain}

@@

\section{Introduction}
\input{body/intro}
\input{body/overview}
\input{body/rbprm}
\input{body/contact}
\input{body/equilibrium}
\input{body/efort}
\input{body/results}
\input{body/conclusion}
\theendnotes
\bibliographystyle{SageH}
\bibliography{biblio}
\appendix
\input{body/appendix_roms}
\input{body/appendix_algo}
\input{body/appendix_hpp}
\end{document}
